---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r library}
library(tidyverse)
```

```{r setup-data}
n <- 1000
set.seed(8675309)
dat <- tibble(
  id = 1:n,
  PSA = rexp(n, 2),
  famhx = rbinom(n, size = 1, prob = .05),
  race = rbinom(n, size = 1, prob = .4)
) %>%
  add_column(
    cancer = rbinom(
      n,
      size = 1,
      prob = 1 / (1 + exp(2 - .$PSA - .$famhx - .$race))
    )
  )
```

### Log binomial fit

```{r fit-log-binom-glm, error=TRUE}
glm(cancer ~ PSA + famhx + race, data = dat, family = binomial(link = "log"))
```

### Poisson fit

```{r fit-poisson}
fit2 <- glm(cancer ~ PSA + famhx + race, data = dat, family = poisson)
fit2
```

### Poisson fit with lme4

```{r fit-poisson-lme4}
fit3 <- lme4::glmer(cancer ~ PSA + famhx + race + (1 | id),
  data = dat, family = poisson
)
fit3
```

### Log binomial fit with starting values from Poisson

In theory, this should work with `coef(fit2)` in the start param, but that throws an error because these coefficients are either outside of or too close to the feasible region boundary where coefficients are bound to `[0, 1]`.

```{r log-binom-start-poisson-error, error = TRUE}
coef(fit2)
fit4 <- glm(cancer ~ PSA + famhx + race,
  data = dat,
  family = binomial(link = "log"),
  start = coef(fit2)
)
```

## negLogLik

Per a suggestion from 
[stats.stackexchange](https://stats.stackexchange.com/questions/105633/what-to-do-when-a-log-binomial-models-convergence-fails),
but replacing 1 in the `pmin()` function with .9999 to avoid `NA`/`Inf` warnings.

```{r negLogLik}
# https://stats.stackexchange.com/a/321407/20236
negLogLik <- function(b) {
  risk <- pmin(.9999, exp(as.matrix(cbind(1, dat[, 2:4])) %*% b))
  -sum(dbinom(dat$cancer, 1, risk, log = TRUE))
}
(fit <- nlm(negLogLik, p = c(log(mean(dat$cancer)), 0, 0, 0), hessian = TRUE))
```

but careful, it's pretty dependent on starting values

```{r negLogLik-2}
(fit <- nlm(negLogLik, p = coef(fit2), hessian = TRUE))
```

fiddling with tolerances may help (but not much the way it's written below)

```{r negLogLik-3}
(fit <- nlm(negLogLik, p = coef(fit2), hessian = TRUE, gradtol = 1e-12, steptol = 1e-11))
(fit <- nlm(negLogLik, p = c(log(mean(dat$cancer)), 0, 0, 0), hessian = TRUE, gradtol = 1e-12, steptol = 1e-11))
```

next steps: can swapping nlm() for optim() help?


## Logbin

```{r logbin}
library(logbin)
(fit <- logbin(cancer ~ PSA + famhx + race, data = dat))
```

It doesn't compute cov matrix to enable CI estimation!

We could use inverted test-based limits 
[according to Agresti](http://statmath.wu.ac.at/research/talks/resources/slidesagresti_confidence.pdf) 
and then simulate coverage review of test-based CIs.

## Example from Spiegelman and Hertzmark AJE 2005

```{r example-spiegelman-aje-2005}
dat <- tibble(
  id = 1:192,
  death = c(rep(1, 54), rep(0, 138)),
  stage = c(
    rep("Stage I", 7), rep("Stage II", 26), rep("Stage III", 21),
    rep("Stage I", 60), rep("Stage II", 70), rep("Stage III", 8)
  ),
  receptor = c(
    rep("Low", 2), rep("High", 5), rep("Low", 9), rep("High", 17),
    rep("Low", 12), rep("High", 9), rep("Low", 10), rep("High", 50),
    rep("Low", 13), rep("High", 57), rep("Low", 2), rep("High", 6)
  )
)
dat
```

log binomial fails to converge

```{r ex-spiegelman-log-bin-fails, error = TRUE}
glm(death ~ stage + receptor, data = dat, family = binomial(link = "log"))
```

Poisson converges

```{r ex-spiegelman-poisson}
(fit <- glm(death ~ stage + receptor, data = dat, family = poisson))
exp(coef(fit)) #' estimates match https://academic.oup.com/aje/article/162/3/199/171116
```

lme4 Poisson

```{r ex-spiegelman-poisson-lme4}
(fit3 <- lme4::glmer(death ~ as.factor(stage) + receptor + (1 | id), data = dat, family = poisson))
```

using Poisson as starting values

```{r ex-spiegelman-poisson-start}
(fit2 <- glm(death ~ as.factor(stage) + receptor, data = dat, family = binomial(link = "log"), start = c(-2.3, .92, 1.78, .489)))
exp(coef(fit2))
```

